Final Project Code ALL

### 2022-11-21 Final_Project_Classification_Models.Rmd
Updated Commits
created combined code file and updated other files (errors remain in classification model)

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(revenue))

@@

# Recipe
data_rec <- recipe(flop ~ No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre, popularity, vote_average, vote_count,
                   data = imdb_clean_class) %>%
  step_naomit()

@@

@@

>>>>

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(revenue - budget))

@@

# Recipe
data_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Meta_score + Genre_1,
                   data = imdb_clean_class) %>%
  step_naomit(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)

@@

We underwent a random forests classification model utilizing OOB evaluation metrics. We also utilized a logistic regression model to create hard predictions.

@@

Included in the above code is OOB importance measures and logistic regression test cases.

### 2022-11-21 Final_Project_Code_ALL.Rmd
created combined code file and updated other files (errors remain in classification model)

---
title: 'STAT 253 Final Project'
author: "Jacob, Julian, Aristo, and Emily"
date: "October 17, 2022"
output: 
  html_document:
    df_print: paged
    code_download: true
---

@@

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(tidymodels)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
```

@@

```{r}
library(readr)
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000.csv")
```

@@

```{r}
imdb_clean <- imdb_top_1000 %>%
  select(-Poster_Link, -Certificate) %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Gross))

@@

imdb_clean$Runtime <- runtime_clean
```

@@

```{r}
full_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>% #variables causing issues
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross)

@@

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

@@

```{r}
spline_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>% #variables causing issues
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6)

@@

The goal of our investigations is to accurately predict a movie's gross box office revenue based on a number of predictors. We compared three different model types: ordinary linear regression (OLS) with all predictors, OLS using LASSO to select a subset of variables, and a GAM using splines to account for nonlinearity. The GAM model has the lowest RMSE but could be difficult to interpret. We are looking forward to continuing our investigations using Decision Trees which may offer a more easily interpreted method for modeling nonlinearity. Ideally, we would like an accurate model that is relatively easy to interpret.

@@

>>>>

---
title: "Final Project Code ALL"
author: "Emily, Julian, Jacob, and Aristo"
date: '2022-11-20'
output: 
  html_document:
    df_print: paged
    code_download: true
---

@@

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(tidymodels)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
library(probably) #install.packages('probably')
library(vip)
```

@@

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

@@

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget)) %>%
  select(-...15)

@@

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf")
```

@@

```{r}
full_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross)

@@

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

@@

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6)

@@

The goal of our investigations is to accurately predict a movie's gross box office gross revenue based on a number of predictors. We compared three different model types: ordinary linear regression (OLS) with all predictors, OLS using LASSO to select a subset of variables, and a GAM using splines to account for nonlinearity. The GAM model has the lowest RMSE but could be difficult to interpret. We are looking forward to continuing our investigations using Decision Trees which may offer a more easily interpreted method for modeling nonlinearity. Ideally, we would like an accurate model that is relatively easy to interpret.

@@

```{r}
imdb_clean_class <- imdb_clean %>%
  mutate(success_ratio = Revenue/Budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 1.5, 'FALSE', 'TRUE'))) %>%
  filter(!is.na(flop)) %>%
  filter(Gross != "NaN") %>%
  filter(success_ratio != "Inf") %>%
  filter(Budget > 10)
```

# Bagging and Random Forests

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = 4, # size of random subset of variables; default is floor(sqrt(number of total predictors))
           trees = 1000, # Number of trees
           min_n = NULL,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') # change this for regression

# Recipe
data_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Meta_score + Genre_1,
                   data = imdb_clean_class) %>%
  step_naomit(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)

# Workflows
data_wf_mtry150 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 150)) %>%
  add_recipe(data_rec)

## Create workflows for min_n = 10, 25, and 50

data_wf_mtry125 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 125)) %>%
  add_recipe(data_rec)

data_wf_mtry75 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 75)) %>%
  add_recipe(data_rec)

data_wf_mtry100 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 100)) %>%
  add_recipe(data_rec)
```

```{r}
# Fit Models
set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry150 <- fit(data_wf_mtry150, data = imdb_clean_class)

# Fit models for 10, 25, and 50
set.seed(123) 
data_fit_mtry125 <- fit(data_wf_mtry125, data = imdb_clean_class)

set.seed(123)
data_fit_mtry75 <- fit(data_wf_mtry75, data = imdb_clean_class)

set.seed(123) 
data_fit_mtry100 <- fit(data_wf_mtry100, data = imdb_clean_class)
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          flop = truth,
          label = model_label
      )
}

rf_OOB_output(data_fit_mtry150,150, imdb_clean_class %>% pull(flop))
```

```{r}
# Evaluate OOB Metrics

data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry150,150, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry125,125, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry75,75, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry100,100, imdb_clean_class %>% pull(flop))
)

data_rf_OOB_output %>% 
    group_by(label) %>%
    accuracy(truth = flop, estimate = .pred_class)
```

### Evaluating the Forest

```{r}
rf_OOB_output(data_fit_mtry125,125, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = class, estimate= .pred_class)
```

### Variable Importance

```{r}
# Impurity

model_output <-data_fit_mtry125 %>% 
    extract_fit_engine()

model_output %>% 
    vip(num_features = 10) + theme_classic() #based on impurity, 10 meaning the top 10

model_output %>% vip::vi() %>% head()
model_output %>% vip::vi() %>% tail()
```

```{r}
# Permutation

model_output2 <- data_wf_mtry125 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = imdb_clean_class) %>% 
    extract_fit_engine()

model_output2 %>% 
    vip(num_features = 10) + theme_classic()

model_output2 %>% vip::vi() %>% head()
model_output2 %>% vip::vi() %>% tail()
```

# Logistic Regression

```{r}
set.seed(123)

# Logistic Regression Model Spec
logistic_spec <- logistic_reg() %>%
    set_engine('glm') %>%
    set_mode('classification')
# Recipe
logistic_rec <- recipe(flop ~ IMDB_Rating, No_of_Votes, gross,
                   data = imdb_clean_class)
# Workflow (Recipe + Model)
log_wf <- workflow() %>%
    add_recipe(logistic_rec) %>%
    add_model(logistic_spec)
# Fit Model
log_fit <- fit(log_wf, data = imdb_clean_class)

tidy(log_fit)
```

```{r}
log_fit %>% tidy() %>%
  mutate(OR = exp(estimate))
```
```{r}
predict(log_fit, new_data = data.frame(IMDB_Rating = 5, No_of_Votes = 500000, gross = 18), type = "prob"
)
```

```{r}
predict(log_fit, new_data = data.frame(IMDB_Rating = 5, No_of_Votes = 500000, gross = 18), type = "class"
)
```
